---
title: "vignettes"
output: html_document
date: "2023-11-26"
---

```{r setup, include=FALSE}
library(functions)
```

## Demonstrate `fn_cos()`

```{r plot, fig.width=8, fig.height=5, echo=FALSE}
x_values <- seq(0, 10, length.out = 100)

plot(x_values, cos(x_values), type = "p", col = "black", pch = 20, xlab = "x", ylab = "Function Output", main = "fn_sin vs sin(x)", , ylim = c(-3, 1))

k_values <- c(1, 3, 5, 7, 9)
colors <- c("blue", "red", "green", "purple", "orange")

for (i in seq_along(k_values)) {
  lines(x_values, fn_cos(x_values, k_values[i]), col = colors[i], lwd = 2)
}

legend("topright", legend = paste("k =", k_values), col = colors, lwd = 2)


```

## Demonstrate `fn_sin()`

```{r plot, fig.width=8, fig.height=5, echo=FALSE}
x_values <- seq(0, 10, length.out = 100)

plot(x_values, sin(x_values), type = "p", col = "black", pch = 20, xlab = "x", ylab = "Function Output", main = "fn_sin vs sin(x)", , ylim = c(-9, 1))

k_values <- c(1, 3, 5, 7, 9)
colors <- c("blue", "red", "green", "purple", "orange")

for (i in seq_along(k_values)) {
  lines(x_values, fn_sin(x_values, k_values[i]), col = colors[i], lwd = 2)
}

legend("topright", legend = paste("k =", k_values), col = colors, lwd = 2)


```

## Demonstrate `calculate_CI()`

### Data Discription

By analysing the data, we can learn about how air transport change in these years.

The data this week comes from [Eurocontrol](https://ansperformance.eu/data/). The data including traffic reports that count the daily number of flights by country, by airport, and by operator. The portal also offers bulk datasets on topics such as airport traffic, flight efficiency, estimated CO2 emissions, and more.

Here's the [data directory](https://github.com/rfordatascience/tidytuesday/tree/master/data/2022/2022-07-12#data-dictionary).

### Data Analysis

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-07-12/flights.csv')
if (!file.exists("data")) {
  dir.create("data")
  saveRDS(df, file = "data/your_data.rds")
}
data <- data %>% drop_na(MONTH_MON)
year1 <- year(data$FLT_DATE)
data <- mutate(data, year1)
month1 <- month(data$FLT_DATE)
data <- mutate(data, month1)
year_trend <- data %>% group_by(year1) %>% summarise(n = sum(FLT_TOT_1))
```

```{r}
print(ggplot(year_trend, aes(x = year1, y = n)) +
  geom_line() +
  labs(title = "Anual number of flights", x = "Year", y = "Number of flights") +
  theme_minimal()
)

```

The graph shows a huge decline in 2020 due to pandemic. We want to have a look on numbers of flight on each months in 2022.

```{r}
trend_2022 <- data %>% filter(year1 == 2022) %>% group_by(month1) %>% summarise(n = sum(FLT_TOT_1))
ggplot(trend_2022, aes(x = month1, y = n)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(title = "Monthly number of flights", x = "Month", y = "Number of flights") +
  theme_minimal()
```

It seems number of flights increased in 2022. But we still want to know whether it returned to the levels before pandemic. So we choose May in 2022 to compare with May in 2019.

```{r}
May_2022 <- data %>% filter(year1 == 2022) %>% filter(month1 == 5) %>% select(FLT_TOT_1)
calculate_CI(May_2022$FLT_TOT_1)
May_2019 <- data %>% filter(year1 == 2019) %>% filter(month1 == 5) %>% select(FLT_TOT_1)
calculate_CI(May_2019$FLT_TOT_1)

category_names <- c("May_2019", "May_2022")
values_list <- list(May_2019, May_2022)

comparison <- data.frame(
  category = rep(category_names),
  values = unlist(values_list)
)

```

```{r}
ggplot(comparison, aes(x = category, y = values)) +
  geom_boxplot(fill = "lightblue", color = "black", width = 0.6) +
  labs(title = "Comparison between 2019 May and 2022 May", x = "Month", y = "Number of flights") +
  theme_minimal()
```

The boxpot of two months seems similar, we could use t-test to test whether they have statistically difference.

```{r}
t.test(May_2019$FLT_TOT_1, May_2022$FLT_TOT_1)
```

### Conclusion

We could reject the null hypothesis that they have same mean, and conclude that the number of flights at 2019 May is statistically significant higher than at 2022 May. This indicates that it didn't return to the levels before pandemic.

### Functions

Functions being used:

`drop_na()`, `filter()`, `mutate()`, `year()`, `month()`, `group_by()`, `summarize()`, `select()`

`ggplot2`: `geom_line`, `geom_bar`, `geom_boxplot`
